Discrete Quantum Theories and Computing

### 2018-12-18 dissertation.tex
dissertation.lyx "Chapter 1 Introduction" Updating...

After first suggested by Richard Feynman 40 years ago~\cite{Feynman1982Simulating},
IBM, \textitalian{Rigetti}, Google, Microsoft, and many other companies
and countries start to build realistic quantum computers recently
\cite{Shieber2018,Kahn2019}, and seek more efficient ways to simulate
chemical molecules and to break down RSA \cite{boneh1999twenty,wiki:RSAFactoring}.
No matter how to, classical or quantum, attack these two tasks, they
require two different types of quantities. On one hand, the question
and answer of RSA are both discrete integers. On the other hand, even
if the quantities describing molecules, like their positions, are
considered as continuous, we want to argue that it is beyond our ability
to decide whether these quantities are ultimately discrete or continuous.
To better understand this issue without struggling with the quantum
theory, it is instructive to review how classical quantities are characterized
and computed.

@@

Although it is convenient to assume physical quantities is as continuous
as their mathematical models, the physical quantities discussed in
the previous paragraph should not simply represented by real numbers.
Chuang-tzu's stick are made by molecules and cannot be cut in halves
endlessly. The input charge of an integrator must actually be an integer
multiple of the elementary charge. A prefect square cannot be draw
physically. Even we don't draw it, but merely asking whether it exists
in the space is still beyond our ability to answer because whether
the space itself is ultimately discrete or continuous is still an
unsolved question. Even if a prefect square exists physically, we
cannot decide whether it is really a prefect square or an almost prefect
one differing only by an arbitrarily small extent. Similar situation
applying to the length of a slide rule, it is beyond our ability to
decide whether its length is ultimately discrete or continuous. Even
if its length is continuous, identifying which real number represents
its length requires infinite precision, but the precision in reality
is limited generally three or four significant figures~\cite{wiki:AnalogComputer}.
As the results, even if analog computers really store real numbers,
they cannot be read, written, and used to branch the computation based
on their exact values. The opposite of the last statement is sometimes
assumed by real computational models. For example, Blum, Cucker, Shub,
and Smale modeled their BCSS machine to allow the operations of deciding
whether a number is greater or equal to zero over real numbers and
whether a number is exactly zero over complex numbers \cite{BSS1989,Ziegler2007,blum2012complexity}.
Therefore, the computational power of analog computers is far weaker
than the prediction based on this kind of theoretical models.

@@

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{ccYY}
\toprule 
\addlinespace
Physical quantities & Discrete & Discrete with extreme small unit & Continuous?\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational devices & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integers & Real numbers & Real numbers\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How the models realize? & Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computation model by a better discrete model. & 
Not reliably, because precision can never be high enough. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum models &  & Quantum Theories and Computing over Finite Fields & Quantum Interval-Valued Probability Measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison} with their
inspired quantum models. The physical quantity manipulated by the
quantum circuit model, the most widely used model of quantum computing,
is the probability amplitudes of states. \todo{Introduction Cleaned
up until here. }Because we are agnostic about whether probability
amplitude is quantized or can really be valued in the continuum of
complex numbers, we tried to develop two different types of models:
the quantum theories and computing over finite fields are discrete
models, and the quantum interval-valued probability measures incorporate
the idea of measurement precision into the quantum theory. To distinguish
the modified quantum theories from the usual ones, the usual ones
will be called the ``conventional quantum theory'' where necessary
and it will be reviewed in Chapter~\ref{chap:Conventional-Quantum-Theory}~\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}.

\begin{tabularx}{\linewidth}{YYYYYYY}
\toprule 
\addlinespace
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to \ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Expectation value & Defined & Undefined & Undefined & Undefined &  & Defined\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Hadamard transformation & Included & May not be included & Included & Included &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Solve $\usat$ efficiently & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
\todo{Beginning of Merge Left}Since conventional probability amplitude
is in the field of complex numbers, and the field structure is essential
to compute the probability from some probability amplitudes, we in
Chapter~\ref{chap:DQT=000026DQC} tried to build discrete quantum
theories and computing by considering them over finite fields.

@@

>>>>

After first suggested by Richard Feynman 40 years ago~\cite{Feynman1982Simulating},
IBM, \textitalian{Rigetti}, Google, Microsoft, and many other companies
and countries start to build realistic quantum computers recently
\cite{Shieber2018,Kahn2019}, and seek more efficient ways to simulate
chemical molecules and to break down RSA \cite{boneh1999twenty,wiki:RSAFactoring}.
No matter how to, classical or quantum, attack these two tasks, they
require two different types of quantities. On one hand, the question
and answer of RSA are both discrete integers. On the other hand, even
if the quantities describing molecules, like the domain of their position
wave functions or the probability amplitudes of a wave function, are
considered as continuous, we want to argue that these quantities might
later be discovered as discrete. Even if they are ultimately continuous,
this fact may never be proved and the exact value of these quantities
may never be pinned down. To better understand this issue without
struggling with the quantum theory, it is instructive to review how
classical quantities are characterized and computed.

@@

Although it is convenient to assume physical quantities is as continuous
as their mathematical models, each physical quantity discussed in
the previous paragraph could not have a simple one-to-one correspondence
with a real number. Chuang-tzu's stick are made by molecules and cannot
be cut in halves endlessly. The input charge of an integrator must
actually be an integer multiple of the elementary charge. A prefect
square cannot be draw physically. Even we don't draw it, but merely
asking whether it exists in the space is still beyond our ability
to answer because whether the space itself is ultimately discrete
or continuous is still an unsolved question. Even if a prefect square
exists physically, we may not able to decide whether the “square”
we discussed is really prefect or almost prefect differing only by
an arbitrarily small extent. Similar situation applying to the length
of a slide rule, it is beyond our ability to decide whether its length
is ultimately discrete or continuous. Even if its length is continuous,
identifying which real number represents its length requires infinite
precision, but the precision in reality is limited generally three
or four significant figures~\cite{wiki:AnalogComputer}. As the results,
even if analog computers really store real numbers, they cannot be
precisely read, written, and used to branch the computation. The opposite
of the last statement is sometimes assumed by real computational models.
For example, Blum, Cucker, Shub, and Smale modeled their BCSS machine
to allow the operations of deciding whether a number is greater or
equal to zero over real numbers and whether a number is exactly zero
over complex numbers \cite{BSS1989,Ziegler2007,blum2012complexity}.
Therefore, the computational power of analog computers is far weaker
than the prediction based on this kind of theoretical models.

@@

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{ccYY}
\toprule 
\addlinespace
Physical quantities & Discrete & Discrete with extreme small unit & Continuous?\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational devices & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integers & Real numbers & Real numbers\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How the models realize? & Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computation model by a better discrete model. & 
Not reliably, because precision can never be high enough. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum models &  & Quantum Theories and Computing over Finite Fields & Quantum Interval-Valued Probability Measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison} with their
inspired quantum models. Similar to classical analog computers, the
quantum circuit model, the most widely used model of quantum computing,
manipulates a quantity which is not discrete apparently: the probability
amplitudes of quantum states, and it is reasonable to expect the similar
consistence between classical classical quantities and their mathematical
representation applies to the quantum case as well. To distinguish
the modified quantum theories from the usual ones, the usual ones
will be called the ``conventional quantum theory'' where necessary
and it will be reviewed in Chapter~\ref{chap:Conventional-Quantum-Theory}~\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.} On one hand, to address the assumption that the probability amplitudes
are ultimately discrete, we consider the quantum theories and computing
over finite fields because the conventional quantum theory is based
on the field of complex numbers, and it is essential to have a field
structure underlying a quantum theory. On the other hand, if the probability
amplitudes are contiuous, we want to incorporate the idea of measurement
precision into the quantum theory and consider the quantum interval-valued
probability measures.

\todo{Introduction Cleaned up until here. }Because we are agnostic
about whether probability amplitude is quantized or can really be
valued in the continuum of complex numbers, we tried to develop two
different types of models: the quantum theories and computing over
finite fields are discrete models, and the quantum interval-valued
probability measures incorporate the idea of measurement precision
into the quantum theory. .

\todo{Beginning of Merge Left}Since conventional probability amplitude
is in the field of complex numbers, and the field structure is essential
to compute the probability from some probability amplitudes, we in
Chapter~\ref{chap:DQT=000026DQC} tried to build discrete quantum
theories and computing by considering them over finite fields.

@@

\begin{tabularx}{\linewidth}{YYYYYYY}
\toprule 
\addlinespace
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to \ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Expectation value & Defined & Undefined & Undefined & Undefined &  & Defined\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Hadamard transformation & Included & May not be included & Included & Included &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Solve $\usat$ efficiently & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
